---
title: "writing_functions"
author: "Nicole Criscuolo"
date: "2025-10-24"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(rvest)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Start small.

```{r}
x_vec = rnorm(20, mean = 10, sd = 3.5)

(x_vec - mean(x_vec)) / sd(x_vec)
```

Write a function to compute z scores.

```{r}
z_scores = function(x) {
  
  if (!is.numeric(x)) {
    stop("The input x should be numeric")
  }
  
  if (length(x) < 5) {
    stop("Only compute z scores when the input has 5 or more numbers")
  }
  
  z = (x - mean(x)) / sd(x)
  
  z
  
}
```

Let's try our function.

```{r}
z_scores(x = x_vec)

num_vec = rnorm(123, mean = 14, sd = 0.4)

z_scores(x = num_vec)
```

Let's break our function.

```{r, error = TRUE}
z_scores(3)
```

```{r, error = TRUE}
z_scores("my name is Nicole")
```

## Let's compute stuff.

Let's compute and return mean and sd of numeric vector.

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("The input x should be numeric")
  }
  
  if (length(x) < 5) {
    stop("Only compute mean and sd when the input has 5 or more numbers")
  }
  
  mean_x = mean(x, na.rm = TRUE)
  sd_x = sd(x, na.rm = TRUE)
  
  c(mean_x, sd_x)
  
#not necessary just makes output look nicer
  tibble( 
    mean = mean_x,
    sd = sd_x
  )
}
```

```{r}
mean_and_sd(x_vec)
```

Let's *simulate* some data.

```{r}
sim_df =
  tibble(
    x = rnorm(n = 30, mean = 3, sd = 2)
  )

sim_df |> 
  summarize(
    mu_hat = mean(x),
    sigma_hat = sd(x)
  )
```

Write a function to do simulations.

The inputs are 

* `n_subject` is the number of subjects
* `mu` is the true mean
* `sigma~ is the true sd

Function simulates data from a normal and computes sample mean and sd.

```{r, eval = FALSE}
sim_mean_sd = function(n_subj, mu = 3, sigma = 2) { ## default values not necessary, just will have to specify later
  
  sim_df =
    tibble(
      x = rnorm(n = n_subj, mean = mu, sd = sigma)
    )
  
  sim_df |> 
    summarize(
      mu_hat = mean(x),
      sigma_hat = sd(x)
    )

}
```

(we now are sourcing this code chunk)
```{r}
source("sim_mean_sd.R")
```

Let's run this function.

```{r}
sim_mean_sd(n_subj = 3800, mu = 8) ##provide named arguments if default not specified or if it is it overrides it
```

Import LOTR data.

```{r}
fellowship_ring =
  read_excel("data/LotR_Words.xlsx", range = "B3:D6") |> 
  mutate(movie = "Fellowship of the Ring")

two_towers =
  read_excel("data/LotR_Words.xlsx", range = "F3:H6") |> 
  mutate(movie = "Two Towers")

return_king =
  read_excel("data/LotR_Words.xlsx", range = "J3:L6") |> 
  mutate(movie = "Return od the King")

lotr_df =
  bind_rows(fellowship_ring, two_towers, return_king)
```

Turn this into a function

```{r}
lotr_import = function(cell_range, movie_title) {
  
  df =
    read_excel("data/LotR_Words.xlsx", range = cell_range) |> 
    mutate(movie = movie_title)
  
  df
}
```

```{r}
fellowship = lotr_import(cell_range = "B3:D6", movie_title = "Fellowship")
two_towers = lotr_import(cell_range = "F3:H6", movie_title = "Two Towers")
return = lotr_import(cell_range = "J3:L6", movie_title = "Return")

bind_rows(fellowship, two_towers, return)
full_join(fellowship, two_towers)
```


```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

data_marj = 
  nsduh_html |> 
  html_table() |> 
  nth(1) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```

Write an import function.

```{r}
nsduh_import = function(html, table_num) {
  
  table = 
  html |> 
  html_table() |> 
  nth(table_num) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
  table
  
}

nsduh_results = 
  bind_rows(
    nsduh_import(nsduh_html, 1),
    nsduh_import(nsduh_html, 4),
    nsduh_import(nsduh_html, 5)
  )
```





